
%-------------------------------------------------
\section{A Manufactured Solution for the FANS-SA Equations}\label{oliver}
%In this section a set of manufactured solution for the FANS-SA equations is developed.  
To exercise all of the terms in the equation, the
solution must satisfy the no slip wall boundary condition for at least
some portion of the boundary.  To avoid pathological behavior in the
solution and required source terms, we strive to make the manufactured
solution reasonably resemble the inner portion (viscous sublayer +
logarithmic layer) of a zero pressure gradient boundary layer.  To
accomplish this goal, the manufactured solution is built using
well-known correlations for turbulent boundary layers.  The
manufactured solution is developed in
Sections~\ref{sec:velocity_field} through~\ref{sec:sa_state_field}.
Some of the derivatives required to compute the manufactured source
terms are given in Section~\ref{sec:summary_and_derivatives}. 
% The behavior of the solution is shown graphically for particular values of the constant parameters in Section~\ref{sec:soln_plots}.  To check the behavior of the SA model terms, the SA model budget is computed and shown in Section~\ref{sec:sa_budget}.



%-------------------------------------------------
\subsection{Velocity Field} \label{sec:velocity_field}

\subsubsection{Streamwise Velocity}
For the streamwise velocity, we construct the solution by working
backward from the van Driest transformation.  Inverting the adiabatic
wall form of the van Driest transformation~\cite{} gives
%
\begin{equation}
u = \frac{u_{\infty}}{A} \sin \left( \frac{A}{u_{\infty}} u_{eq} \right),
\end{equation}
% 
where $u_{\infty}$, $A = \sqrt{1 - T_{\infty}/T_{aw}}$,
$T_{\infty}$, and $T_{aw}$ are constants, and $u_{eq}$ is the van
Driest equivalent velocity.  The van Driest equivalent velocity can be
computed from the friction velocity $u_{\tau}$ and the
non-dimensionalized van Driest velocity $u_{eq}^+ \equiv
u_{eq}/u_{\tau}$.  Clearly,
%
\begin{equation}
u_{eq} = u_{\tau} u_{eq}^+.
\end{equation}
% 
Thus, to specify the velocity, one must specify $u_{\tau}$ and
$u_{eq}^+$.  The friction velocity can be determined from the skin
friction coefficient:
%
\begin{equation}
u_{\tau} \equiv \sqrt{\frac{\tau_{w}}{\rho_w}} = u_{\infty} \sqrt{\frac{c_f}{2}}.
\end{equation}
% 
We model the skin friction coefficient using the compressibility
transformation idea of Spalding and Chi~\cite{} and a correlation for
the incompressible skin friction.  Specifically,
%
\begin{equation}
c_f = \frac{1}{F_c} c_{f,inc}\left( \frac{1}{F_c} Re_x \right),
\end{equation}
%
where $ F_c = \frac{T_{aw}/T_{\infty} - 1}{ \left( \sin^{-1} A \right)^2} $
%
is a constant and $c_{f,inc}$ is a correlation for the incompressible
skin friction.  Specifically, we choose a power law for the
incompressible skin friction coefficient:
%
\begin{equation}
c_{f,inc}(Re_x) = C_{cf} Re_{x}^{-1/7},
\end{equation}
%
where $C_{cf}$ is a constant.

Finally, to complete the manufactured solution, we set $u^+_{eq}$
using the velocity profile model of Cebeci and Bradshaw~\cite{}:
%
\begin{equation}
u_{eq}^+ = \frac{1}{\kappa} \log \left( 1 + \kappa y^+ \right) + C_1 \left[ 1 - e^{-y^+/\eta_1} - \frac{y^+}{\eta_1} e^{-y^+ b} \right],
\end{equation}
%
where $\kappa$, $C_1 = -(1/\kappa) \log(\kappa) + C$, $C$, $\eta_1$ and $b$ are constants and
%
\begin{equation}
y^+ \equiv \frac{y}{\ell_v}, \quad\text{and}\quad 
\ell_v \equiv \frac{\nu_w}{u_{\tau}}.
\end{equation}
%


\subsubsection{Wall-normal Velocity}
We choose
%
\begin{equation}
v = -\eta_v \dd{u_{\tau}}{x} y
\end{equation}
%
where $\eta_v$ is a user-specified parameter.

%-------------------------------------------------
\subsection{Thermodynamic State} \label{sec:thermodynamic_state}
For the mean temperature, we write
%
\begin{equation}
T = T_{\infty} \left[ 1 + r \frac{\gamma - 1}{2} M_{\infty}^2 \left( 1 - \left(\frac{u}{u_{\infty}}\right)^2 \right) \right],
\end{equation}
% 
where $T_{\infty}$, $r$, and $\gamma$ are additional constant
parameters.  Note that the constant $T_{aw}$ is defined as follows:
%
\begin{equation}
T_{aw} = T(u=0) = T_{\infty} \left[ 1 + r \frac{\gamma - 1}{2} M_{\infty}^2 \right].
\end{equation}
%

Choosing the pressure to be a constant $p = p_0$, the
density can be computed from the ideal gas equation:
%
\begin{equation}
\rho = \frac{p_0}{R T},
\end{equation}
%
where $R$ is the gas constant.

%-------------------------------------------------
\subsection{Spalart-Allmaras Variable} \label{sec:sa_state_field}
The velocity profile manufactured solution constructed above is
intended to be a reasonable representation of the viscous sublayer and
log layer in a zero pressure gradient boundary layer.  In this region
in an incompressible boundary layer, the SA model is designed to give
$\sa/\nu = \kappa y^+$.  Based on this form, we choose
%
\begin{equation}
\sa = \kappa u_{\tau} y - \alpha y^2,
\end{equation}
%
where $\alpha$ is a constant.  The $y^2$ term is included simply to
make the solution nonlinear in $y$.

%-------------------------------------------------
\subsection{Manufactured Solution Summary and Spatial Derivatives} \label{sec:summary_and_derivatives}

\subsubsection{Streamwise Velocity}
The streamwise velocity field is defined by
%
\begin{equation*}
u = \frac{u_{\infty}}{A} \sin \left( \frac{A}{u_{\infty}} u_{eq} \right).
\end{equation*}
%
The values $u_{\infty}$ and $A$ are constants.  Thus,
%
\begin{gather*}
\pp{u}{x} = \cos \left(\frac{A}{u_{\infty}} u_{eq} \right) \pp{u_{eq}}{x}, \quad \text{and} \quad
\pp{u}{y} = \cos \left(\frac{A}{u_{\infty}} u_{eq} \right) \pp{u_{eq}}{y}.
\end{gather*}
%

The van Driest equivalent velocity is given by
%
\begin{equation*}
u_{eq} = u_{\tau} u_{eq}^+,
\end{equation*}
% 
where $u_{\tau}$ is only a function of $x$ and $u_{eq}^+$ is only a
function of $y^+ \equiv (y u_{\tau})/\nu_w$.  Thus,
%
\begin{gather*}
\pp{u_{eq}}{x} = \dd{u_{\tau}}{x} u_{eq}^+ + u_{\tau} \dd{u_{eq}^+}{y^+} \pp{y^+}{x} \quad \text{and} \quad
\pp{u_{eq}}{y} =  u_{\tau} \dd{u_{eq}^+}{y^+} \pp{y^+}{y},
\end{gather*}
%
where
%
\begin{gather*}
\pp{y^+}{x} = \frac{y}{\nu_w} \dd{u_{\tau}}{x}, \quad \text{and} \quad
\pp{y^+}{y} = \frac{u_{\tau}}{\nu_w}.
\end{gather*}
%

The friction velocity is given by
%
\begin{equation*}
u_{\tau} = u_{\infty} \sqrt{\frac{c_f}{2}}.
\end{equation*}
%
Thus,
%
\begin{equation*}
\dd{u_{\tau}}{x} = u_{\infty}  \frac{1}{2} \sqrt{ \frac{2}{c_f} } \dd{c_f}{x}.
\end{equation*}
%

The skin friction coefficient is given by
%
\begin{equation*}
c_f = \frac{C_{cf}}{F_c} \left( \frac{1}{F_c} Re_x \right)^{-1/7} = \frac{C_{cf}}{F_c} \left( \frac{1}{F_c} \frac{\rho_{\infty} u_{\infty} x}{\mu} \right)^{-1/7}
\end{equation*}
%
where $C_{cf}$, $F_c$, $\rho_{\infty}$, $u_{\infty}$, and $\mu$ are constants.  Thus,
%
\begin{equation*}
\dd{c_f}{x} = - \, \frac{C_{cf}}{F_c} \frac{1}{7} \left( \frac{1}{F_c} Re_x \right)^{-8/7} \frac{1}{F_c} \frac{\rho_{\infty} u_{\infty}}{\mu}
\end{equation*}

Finally, the non-dimensionalized van Driest velocity profile is given
by
%
\begin{equation*}
u_{eq}^+ = \frac{1}{\kappa} \log \left( 1 + \kappa y^+ \right) + C_1 \left[ 1 - e^{-y^+/\eta_1} - \frac{y^+}{\eta_1} e^{-y^+ b} \right].
\end{equation*}
%
Thus,
%
\begin{equation*}
\dd{u_{eq}^+}{y^+} = \frac{1}{\left( 1 + \kappa y^+ \right)} + C_1 \left[ \frac{1}{\eta_1} e^{-y^+/\eta_1} - \frac{1}{\eta_1} e^{-y^+ b} + b \frac{y^+}{\eta_1} e^{-y^+ b} \right].
\end{equation*}
%

\subsubsection{Wall-normal Velocity}
The wall-normal velocity is given by
%
\begin{equation*}
v = -\eta_v \dd{u_{\tau}}{x} y.
\end{equation*}
%
Thus,
%
\begin{gather*}
\pp{v}{x} = - \eta_v \dd{^2 u_{\tau}}{x^2} y, \quad \text{and} \quad
\pp{v}{y} = - \eta_v \dd{u_{\tau}}{x}.
\end{gather*}
%

\subsubsection{Thermodynamic Variables}
The temperature is given by
%
\begin{equation*}
T = T_{\infty} \left[ 1 + r \frac{\gamma - 1}{2} M_{\infty}^2 \left( 1 - \left(\frac{u}{u_{\infty}}\right)^2 \right) \right].
\end{equation*}
%
Thus,
%
\begin{gather*}
\pp{T}{x} = T_{\infty} \left[  r \frac{\gamma - 1}{2} M_{\infty}^2 \left(  - 2 \left(\frac{u}{u_{\infty}}\right) \frac{1}{u_{\infty}} \pp{u}{x} \right) \right], \quad \text{and} \quad
\pp{T}{y} = T_{\infty} \left[  r \frac{\gamma - 1}{2} M_{\infty}^2 \left(  - 2 \left(\frac{u}{u_{\infty}}\right) \frac{1}{u_{\infty}} \pp{u}{y} \right) \right].
\end{gather*}
%

The density is given by
%
\begin{equation*}
\rho = \frac{p_0}{R T}.
\end{equation*}
%
Thus,
%
\begin{gather*}
\pp{\rho}{x} = - \, \frac{p_0}{R T^2} \pp{T}{x}, \quad \text{and} \quad
\pp{\rho}{y} = - \, \frac{p_0}{R T^2} \pp{T}{y}. 
\end{gather*}

\subsubsection{Spalart-Allmaras Variable}
The SA state variable is given by
%
\begin{equation*}
\sa = \kappa u_{\tau} y - \alpha y^2.
\end{equation*}
%
Thus,
%
\begin{gather*}
\pp{\sa}{x} = \kappa \dd{u_{\tau}}{x} y \quad \text{and} \quad
\pp{\sa}{y} = \kappa u_{\tau} - 2 \alpha y.
\end{gather*}
%
